------------------------------------------------------------------------

---
title: "Data Analysis of Reframing Pavlovian Bias: Semantic Framing Reshapes Action-Valence Associations in a Go/NoGo Task"
format: html
editor: visual
---

# **Data analysis of \[TITLE\]**

```{r}
# Required packages
if (!require(dplyr, quietly = TRUE)) {install.packages("dplyr")}
options(dplyr.summarise.inform=FALSE)
if (!require(tidyr, quietly = TRUE)) {install.packages("tidyr")} 
if (!require(lattice, quietly = TRUE)) {install.packages("lattice")}
if (!require(ggplot2, quietly = TRUE)) {install.packages("ggplot2")}
if (!require(ggrain, quietly = TRUE)) {install.packages("ggrain")}
if (!require(ggpubr, quietly = TRUE)) {install.packages("ggpubr")}
if (!require(carData, quietly = TRUE)) {install.packages("carData")}
if (!require(brms, quietly = TRUE)) {install.packages("brms")}
if (!require(ggeffects, quietly = TRUE)) {install.packages("ggeffects")}

```

## Data, preprocessing, filtering & demographics

```{r}
df_data <- read.csv("https://raw.githubusercontent.com/JelleGoetschalckx/RPEP/refs/heads/main/data_all.csv")

df_data$outcome[df_data$outcome == "punishment"] <- "loss"
df_data$framing <- factor(df_data$framing)
df_data$shape <- factor(df_data$shape)
df_data$color <- factor(df_data$color)
df_data$outcome <- factor(df_data$outcome)
df_data$correct_response <- factor(df_data$correct_response)
df_data$given_response <- factor(df_data$given_response)
df_data$feedback <- factor(df_data$feedback)
df_data$gender <- factor(df_data$gender)
df_data$colorblind <- factor(df_data$colorblind)
df_data$first_framing <- factor(df_data$first_framing)

options(contrasts = c("contr.sum", "contr.poly"))
df_data$outcome <- relevel(df_data$outcome, "reward")
df_data$framing <- relevel(df_data$framing, "congruent")
df_data$correct_response <- relevel(df_data$correct_response, "Go")

```

### Exclusion criteria

#### Exclusion criterion 1: removal of participants with an accuracy 3 SD lower than mean

```{r}
# Accuracy 3 standard deviations from mean
exclusion_crit_1 <- df_data |>
  group_by(nr) |>
  summarise(
    mean = mean(accuracy),
    sd = sd(accuracy)
  ) |>
  ungroup() |>
  mutate(
    group_sd = sd(mean),
    group_mean = mean(mean)
  ) |>
  mutate(
    z = (mean - group_mean) / group_sd
  ) |> 
  filter(z < -3)

```

```{r}
# Remove these participants from data
df_data <- df_data |>
  filter(!(nr %in% (exclusion_crit_1)$nr))

# Participants 42 & 103 removed (total of 640 trials)

n_trials <- 320
n <- as.numeric(nrow(df_data)) / n_trials

```

#### Exclusion criterion 2: same response on 90% of trials

```{r}
exclusion_crit_2 <- data.frame((
  df_data |>
    group_by(nr) |>
    count(given_response) |>
    mutate(n = n / n_trials) |>
    filter(n > 0.9 | n < 0.1)
)$nr)

### None found

```

#### Exclusion criterion 3: implausibly fast reaction times (below 200 ms)

```{r}
exclusion_crit_3 <- df_data |>
  filter(rt<=0.2)
cat("Participant with the highest rate of too fast responses answered implausibly fast in", round(max(table(exclusion_crit_3$nr))/(n_trials/2)*100), "% of Go-trials")

cat("\n\nTotal amount of removed trials (response faster than 200ms):", nrow(exclusion_crit_3))

```

```{r}
## Remove trials with responses faster than 200ms

df_data <- df_data |>
  filter(rt>0.2 | is.na(rt))

# 53 trials removed

```

Total amount of removed trials: (640 + 53) = 693

### Demographics

```{r}
# Create dataframe with one row per participant and their summarized data
df_indiv_data <- df_data |>
  group_by(nr, framing) |>
  summarise(
    across(c(age, gender, colorblind, was_first_exp), first), 
    acc = mean(accuracy),
    rt = mean(rt, na.rm = TRUE),
    times_read = first(times_instructions_read)
  ) |>
  pivot_wider(
    id_cols = c(nr, age, gender, colorblind, was_first_exp),
    names_from = framing,
    values_from = c(acc, rt, times_read),
  ) |>
  mutate(
    mean_acc = mean(c(acc_congruent, acc_incongruent)),
    mean_rt = mean(c(rt_congruent, rt_incongruent)),
    age = replace(age, age == 0, NA)
  ) |>
  relocate(
    c(
      nr,
      age,
      gender,
      colorblind,
      acc_congruent,
      acc_incongruent,
      mean_acc,
      rt_congruent,
      rt_incongruent,
      mean_rt,
      was_first_exp
    )
  )

```

```{r}
# Print demographics of participants
cat("Age mean:", mean(df_indiv_data$age, na.rm=TRUE), "| Standard deviation:", sd(df_indiv_data$age, na.rm=TRUE), "\n\n")
cat("Table of reported ages:")
table(df_indiv_data$age, useNA = "always")

cat("\n\nTables of reported gender and proportion (%):")
table(df_indiv_data$gender)
round(prop.table(table(df_indiv_data$gender)) * 100, 2)

cat("\n\n# Colorblind participants:", levels(df_indiv_data$colorblind))

```

### Difference in amount of time the instructions were read between framings

```{r}
mean(df_indiv_data$times_read_congruent)
sd(df_indiv_data$times_read_congruent)

mean(df_indiv_data$times_read_incongruent)
sd(df_indiv_data$times_read_incongruent)

t.test(df_indiv_data$times_read_congruent,
       df_indiv_data$times_read_incongruent,
       paired = TRUE)

# Result: No significant difference
```

### Effect of starting with current experiment or after other experiment (e.g. fatigue)

```{r}
exp_as_1st <- df_indiv_data |> filter(was_first_exp == 1)
cat("_Exp as 1st_\n")
cat("Acc (M, SD)\n")
mean(exp_as_1st$mean_acc)
sd(exp_as_1st$mean_acc)
cat("RT (M, SD)\n")
mean(exp_as_1st$mean_rt)
sd(exp_as_1st$mean_rt)

exp_as_2nd <- df_indiv_data |> filter(was_first_exp == 0)
cat("_Exp as 2nd_\n")
cat("Acc (M, SD)\n")
mean(exp_as_2nd$mean_acc)
sd(exp_as_2nd$mean_acc)
cat("RT (M, SD)\n")
mean(exp_as_2nd$mean_rt)
sd(exp_as_2nd$mean_rt)

t.test(exp_as_1st$mean_acc, exp_as_2nd$mean_acc)

t.test(exp_as_1st$mean_rt, exp_as_2nd$mean_rt)

```

## Plots

```{r}
# Function for plot creation
raincloud_plot <- function(df, df_sd, type) {
  geom_rain_layer <- if (type == "rt") {
    geom_rain(
      alpha = .8,
      rain.side = "l",
      point.args = list(alpha = 0.5),
      boxplot.args.pos = list(width = 0.05, position = position_nudge(x = 0.13)),
      violin.args.pos = list(
        side = "r",
        width = 0.8,
        position = position_nudge(x = 0.2)
      )
    )
  } else {
    geom_rain(
      alpha = .7,
      rain.side = "l",
      point.args = list(alpha = 0.3),
      point.args.pos = list(position = position_jitter(
        width = 0.08,
        height = 0.0035,
        seed = 1
      )), 
      boxplot.args = list(
        color = NA,
        fill = NA
      ),
      violin.args.pos = list(
        side = "r",
        width = 1,
        position = position_nudge(x = 0.15)
      )
    )
  }
  return (
    ggplot(
      df,
      aes(x = response_and_outcome, y = mean, fill = response_and_outcome)
    ) +
      geom_rain_layer +
      stat_summary(
        fun = "mean",
        geom = "point",
        position = position_nudge(x = .2)
      ) +
      geom_errorbar(
        inherit.aes = FALSE,
        df_sd,
        mapping = aes(x = response_and_outcome, ymin = CI_lower, ymax = CI_higher),
        width = 0.1,
        position = position_nudge(x = .2)
    )
  )
}

# Function for cleaning up plot visuals
plot_cleaner <- function(plot, colors, ylims = c(0, 1), label_data) {
  plot +
    coord_cartesian(ylim = c(ylims[1], ylims[2])) +
    labs(fill = NULL) +
    rremove("xlab") + rremove("ylab") +
    theme(
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(colour = "lightgrey"),
      panel.grid.minor = element_line(colour = "lightgrey"),
      panel.grid.major.x = element_line(colour = "white")
    ) +
    scale_fill_manual(values = colors) +
    theme(legend.position = "none") +
    scale_x_discrete(labels = if (length(colors) == 4) {
      c(
        paste(
          "Go\nReward\n",
          "M =", round(label_data[1, 3], 3),
          "\nSD =", round(label_data[1, 2], 3)
        ),
        paste(
          "Go\nLoss\n",
          "M =", round(label_data[2, 3], 3),
          "\nSD =", round(label_data[2, 2], 3)
        ),
        paste(
          "NoGo\nReward\n",
          "M =", round(label_data[3, 3], 3),
          "\nSD =", round(label_data[3, 2], 3)
        ),
        paste(
          "NoGo\nLoss\n",
          "M =", round(label_data[4, 3], 3),
          "\nSD =", round(label_data[4, 2], 3)
        )
      )
    } else {
      c(
        paste(
          "Go Reward\n",
          "M =", round(label_data[1, 3], 3),
          "\nSD =", round(label_data[1, 2], 3)
        ),
        paste(
          "Go Loss\n",
          "M =", round(label_data[2, 3], 3),
          "\nSD =", round(label_data[2, 2], 3)
        )
      )
    })
}

# Function for changing colors of plots
recolor_plot <- function(plot, colors) {
  return(suppressMessages(plot + scale_fill_manual(values = colors)))
}

# Function for placing plots next to eachother
combine_plots <- function(plot1, plot2, type) {
  plot1_labeled <- annotate_figure(
    plot1, 
    top = text_grob("Congruent framing", vjust = ifelse(type=="rt", 2.3, 1.25))
  )
  plot2_labeled <- annotate_figure(
    plot2, 
    top = text_grob("Incongruent framing", vjust = ifelse(type=="rt", 2.3, 1.25))
  )
  plot_combined <- ggarrange(plot1_labeled, plot2_labeled, labels = NULL)
  annotate_figure(
    plot_combined,
    left = if (type == "rt") "Response time (ms)" else "Accuracy (%)"
  )
}

```

### Plot of reaction times

```{r}
# Make dataframes to plot
rt_congruent <- df_data |>
  filter(framing=="congruent", correct_response=="Go") |>
  group_by(nr, outcome, correct_response) |>
  summarise(
    mean = mean(rt, na.rm=TRUE),
    sd = sd(rt, na.rm=TRUE)
  ) |>
  mutate(response_and_outcome = factor(paste(correct_response, outcome) , levels = c("Go reward", "Go loss")))

sd_rt_congruent <- rt_congruent |>
  group_by(response_and_outcome) |>
  summarise(
    sd = sd(mean, na.rm=TRUE),
    total_mean = mean(mean, na.rm=TRUE),
    CI_lower = total_mean - (sd * 1.96)/sqrt(n),
    CI_higher = total_mean + (sd * 1.96)/sqrt(n)
  )

rt_incongruent <- df_data |>
  filter(framing=="incongruent", correct_response=="Go") |>
  group_by(nr, outcome, correct_response) |>
  summarise(
    mean = mean(rt, na.rm=TRUE),
    sd = sd(rt, na.rm=TRUE)
  ) |>
  mutate(response_and_outcome = factor(paste(correct_response, outcome), levels = c("Go reward", "Go loss")))

sd_rt_incongruent <- rt_incongruent |>
  group_by(response_and_outcome) |>
  summarise(
    sd = sd(mean, na.rm=TRUE),
    total_mean = mean(mean, na.rm=TRUE),
    CI_lower = total_mean - (sd * 1.96)/sqrt(n),
    CI_higher = total_mean + (sd * 1.96)/sqrt(n)
  )

```

```{r}
plot_rt_congruent <- raincloud_plot(rt_congruent, sd_rt_congruent, "rt") +
  scale_y_continuous(labels = function(x)x * 1000)

plot_rt_incongruent <- raincloud_plot(rt_incongruent, sd_rt_incongruent, "rt") + 
  scale_y_continuous(labels = function(x)x * 1000)

plot_rt_congruent <- plot_cleaner(
  plot_rt_congruent, 
  colors = c("purple", "blue"), 
  ylims=c(0.28, 0.62),
  label_data = sd_rt_congruent
)
plot_rt_incongruent <- plot_cleaner(
  plot_rt_incongruent, 
  colors = c("yellow", "pink"), 
  ylims=c(0.28, 0.62),
  label_data = sd_rt_incongruent
)

rt_plot <- combine_plots(plot_rt_congruent, plot_rt_incongruent, type="rt")
rt_plot

```

### Plot of accuracy

```{r}
acc_congruent <- df_data |>
  filter(framing == "congruent") |>
  group_by(nr, outcome, correct_response) |>
  summarise(
    mean = mean(accuracy, na.rm = TRUE),
    sd = sd(accuracy, na.rm = TRUE)
  ) |>
  mutate(response_and_outcome = factor(
    paste(correct_response, outcome),
    levels = c("Go reward", "Go loss", "NoGo reward", "NoGo loss")
  ))

sd_acc_congruent <- acc_congruent |>
  group_by(response_and_outcome) |>
  summarise(
    sd = sd(mean, na.rm = TRUE),
    total_mean = mean(mean, na.rm = TRUE),
    CI_lower = total_mean - 1.96 * sd / sqrt(n),
    CI_higher = total_mean + 1.96 * sd / sqrt(n)
  )

acc_incongruent <- df_data |>
  filter(framing == "incongruent") |>
  group_by(nr, outcome, correct_response) |>
  summarise(
    mean = mean(accuracy, na.rm = TRUE),
    sd = sd(accuracy, na.rm = TRUE)
  ) |>
  mutate(response_and_outcome = factor(
    paste(correct_response, outcome),
    levels = c("Go reward", "Go loss", "NoGo reward", "NoGo loss")
  ))

sd_acc_incongruent <- acc_incongruent |>
  group_by(response_and_outcome) |>
  summarise(
    sd = sd(mean, na.rm = TRUE),
    total_mean = mean(mean, na.rm = TRUE),
    CI_lower = total_mean - 1.96 * sd / sqrt(n),
    CI_higher = total_mean + 1.96 * sd / sqrt(n)
  )

```

```{r}
plot_acc_congruent <- raincloud_plot(acc_congruent, sd_acc_congruent, "acc") +
  scale_y_continuous(labels = function(x)x * 100)

plot_acc_incongruent <- raincloud_plot(acc_incongruent, sd_acc_incongruent, "acc") +
  scale_y_continuous(labels = function(x)x * 100)


plot_acc_congruent <- plot_cleaner(
  plot_acc_congruent, 
  colors = c("purple", "blue", "purple", "blue"), 
  ylims=c(0.85, 1),
  label_data = sd_acc_congruent
)
plot_acc_incongruent <- plot_cleaner(
  plot_acc_incongruent, 
  colors = c("yellow", "pink", "yellow", "pink"), 
  ylims=c(0.85, 1),
  label_data = sd_acc_incongruent
)

acc_plot <- combine_plots(plot_acc_congruent, plot_acc_incongruent, type="acc")
acc_plot

```

### Log distribution of RT fits normal distribution better

```{r}
ggqqplot((df_data |> filter(!is.na(rt)))$rt)
ggqqplot(log((df_data |> filter(!is.na(rt)))$rt))

```

## Linear Mixed Models

```{r}
# Uncomment to run, recommended to run chunk below instead
# fit_acc <- afex::mixed(
#   # Fixed effects
#   accuracy ~ framing * correct_response * outcome + (framing * first_framing) +
#   # Random effects
#   (framing * correct_response * outcome | nr),
#   # Data, binomial for binary accuracy
#   data = df_data,
#   family = binomial,
#   method = "LRT",
#   control = lme4::glmerControl(
#     optCtrl = list(maxfun = 100000),
#     calc.derivs = FALSE
#   )
# )

```

```{r}
download.file("https://raw.githubusercontent.com/JelleGoetschalckx/RPEP/main/models/fit_acc_model", destfile = "fit_acc_model.rds", mode = "wb")
fit_acc <- readRDS("fit_acc_model.rds")
summary(fit_acc)

```

```{r}
# Uncomment to run, recommended to run chunk below instead
# fit_rt <- afex::mixed(
#   # Fixed effects (log of rt for more normally distributed rt)
#   log(rt) ~ framing * outcome + (framing * first_framing) +
#   # Random effects
#   (framing * outcome | nr),
#   # Data
#   data = df_data |> filter(correct_response == "Go") |> filter(!is.na(rt)),
#   # Settings
#   control = lme4::lmerControl(
#     optimizer = "bobyqa",
#     optCtrl = list(maxfun = 100000),
#     calc.derivs = FALSE
#   )
# )

```

```{r}
download.file("https://raw.githubusercontent.com/JelleGoetschalckx/RPEP/main/models/fit_rt_model", destfile = "fit_rt_model.rds", mode = "wb")
fit_rt <- readRDS("fit_rt_model.rds")
summary(fit_rt)

```

Below, hypotheses are tested. The used contrasts can be found below, hypothesis-irrelevant factors were always left out by multiplying it with 0.

```{r}
contrasts(df_data$framing)
contrasts(df_data$correct_response)
contrasts(df_data$outcome)
contrasts(df_data$first_framing)

```

### Hypothesis 1: Replication of Pavlovian Biases in the Congruent Framing

1.1 G-R trials will yield more accurate responses than G-L trials.

```{r}
# Accuracy G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# correct_response            1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response    1           1           | 0
# framing:outcome             1          -1           | 2
# correct_response:outcome    1          -1           | 2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   1          -1           | 2

L_h1.1 <- c(0, 0, 0, 2, 0, 0, 2, 2, 0, 2)

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h1.1)

# Plot
recolor_plot(plot_acc_congruent, c("red", "red", "white", "white"))

# Result: significant, reflects Pavlovian bias

```

1.2 G-R trials will yield faster response times than G-L trials.

```{r}
# RT G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:outcome             1          -1           | 2
# framing:first_framing       0           0           | 0

L_h1.2 <- c(0, 0, 2, 0, 2, 0)

# Hypothesis test
car::linearHypothesis(fit_rt$full_model, L_h1.2, test = "F")

# Plot
combine_plots(
  recolor_plot(plot_rt_congruent, c("red", "red")),
  recolor_plot(plot_rt_incongruent, c("white", "white")),
  type = "rt"
)

# Result: significant, reflects Pavlovian bias

```

1.3 NG-L trials will yield more accurate responses than NG-R trials.

```{r}
# Accuracy NG-R vs. NoGo L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# correct_response           -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response   -1          -1           | 0
# framing:outcome             1          -1           | 2
# correct_response:outcome   -1           1           |-2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -1           1           |-2

L_h1.3 <- c(0, 0, 0, 2, 0, 0, 2, -2, 0, -2)

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h1.3)

# Plot
recolor_plot(plot_acc_congruent, c("white", "white", "red", "red"))

# Result: significant, reflects Pavlovian bias

```

### Hypothesis 2: Collapse of Pavlovian Bias in the Incongruent Framing

2.1 G-R trials will yield equally or less accurate responses compared to G-L trials.

```{r}
# Accuracy G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# correct_response            1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response   -1          -1           | 0
# framing:outcome            -1           1           |-2
# correct_response:outcome    1          -1           | 2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -1           1           |-2

L_h2.1 <- c(0, 0, 0, 2, 0, 0, -2, 2, 0, -2)

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h2.1)

# Plot
recolor_plot(plot_acc_incongruent, c("red", "red", "white", "white"))

# Result: not significant, reflects collapse of Pavlovian bias

```

2.2 G-R trials will yield equally or slower response times compared to G-L trials.

```{r}
# RT G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:outcome            -1           1           |-2
# framing:first_framing       0           0           | 0

L_h2.2 <- c(0, 0, 2, 0, -2, 0)

# Hypothesis test
car::linearHypothesis(fit_rt$full_model, L_h2.2, test = "F")

# Plot
combine_plots(
  recolor_plot(plot_rt_congruent, c("white", "white")),
  recolor_plot(plot_rt_incongruent, c("red", "red")),
  type = "rt"
)

# Result: not significant, reflects collapse of Pavlovian bias

```

2.3 NG-R trials will yield equally or less accurate responses compared to NG-L trials.

```{r}
# Accuracy NG-R vs. NG-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# correct_response           -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response    1           1           | 0
# framing:outcome            -1           1           |-2
# correct_response:outcome   -1           1           |-2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   1          -1           | 2

L_h2.3 <- c(0, 0, 0, 2, 0, 0, -2, -2, 0, 2)

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h2.3)

# Plot
recolor_plot(plot_acc_incongruent, c("white", "white", "red", "red"))

# Result: significant, but effect goes the other way; reflects opposite of Pavlovian bias

```

### Hypothesis 3: Significant difference in Pavlovian bias effects

3.1 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for accuracy in G-R vs. G-L trials.

```{r}
# Accuracy Congruent vs. Incongruent for G-R vs. G-L
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# correct_response            0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:correct_response    0           0           | 0
# framing:outcome             2          -2           | 4
# correct_response:outcome    2           2           | 0
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   2          -2           | 4

L_h3.1 <- L_h1.1 - L_h2.1

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h3.1)

# Plots
combine_plots(
  recolor_plot(plot_acc_congruent, c("brown", "brown", "white", "white")),
  recolor_plot(plot_acc_incongruent, c("red", "red", "white", "white")),
  type = "acc"
)

# Result: significant difference in Pavlovian bias effect, indicating collapse

```

3.2 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for response times in G-R vs. G-L trials.

```{r}
# RT Congruent vs. Incongruent for G-R vs. G-L
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:outcome             2          -2           | 4
# framing:first_framing       0           0           | 0

L_h3.2 <- L_h1.2 - L_h2.2

# Hypothesis test
car::linearHypothesis(fit_rt$full_model, L_h3.2)

# Plots
combine_plots(
  recolor_plot(plot_rt_congruent, c("brown", "brown")),
  recolor_plot(plot_rt_incongruent, c("red", "red")),
  type = "rt"
)

# Result: significant difference in Pavlovian bias effect, indicating collapse


```

3.3 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for accuracy in NG-R vs. NG-L trials.

```{r}
# Accuracy Congruent vs. Incongruent for NG-R vs. NG-L
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# correct_response            0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:correct_response    0           0           | 0
# framing:outcome             2          -2           | 4
# correct_response:outcome   -2          -2           | 0
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -2           2           |-4

L_h3.3 <- L_h1.3 - L_h2.3

# Hypothesis test
car::linearHypothesis(fit_acc$full_model, L_h3.3)

# Plots
combine_plots(
  recolor_plot(plot_acc_congruent, c("white", "white", "brown", "brown")),
  recolor_plot(plot_acc_incongruent, c("white", "white", "red", "red")),
  type = "acc"
)

# Result: !!

```

### Hypothesis 4: Interaction effect framing and first framing

```{r}
L_h4_acc <- c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
car::linearHypothesis(fit_acc$full_model, L_h4_acc)

L_h4_rt <- c(0, 0, 0, 0, 0, 1)
car::linearHypothesis(fit_rt$full_model, L_h4_rt)

# Significant effect, but this is likely due to practice effects. This is visible in the plots below.

```

```{r}
interaction.plot(
  x.factor = factor(df_data$framing==df_data$first_framing, levels=c(TRUE, FALSE)),
  xlab = "",
  trace.factor = df_data$first_framing,
  trace.label = "First framing",
  response = df_data$accuracy,
  ylab = "Accuracy (%)",
  col = c("pink", "blue"),
  lty = 1,
  lwd = 2,
  xaxt = "n"
)
mtext("Time", side = 1, line = 1)

interaction.plot(
  x.factor = df_data$trial_nr,
  xlab = "",
  trace.factor = df_data$first_framing,
  trace.label = "First framing",
  response = df_data$accuracy,
  ylab = "Accuracy (%)",
  col = c("pink", "blue"),
  lty = 1,
  lwd = 2,
  xaxt = "n"
)
mtext("Time", side = 1, line = 1)

interaction.plot(
  x.factor = factor((df_data|>filter(!is.na(rt)))$framing==(df_data|>filter(!is.na(rt)))$first_framing, levels=c(TRUE, FALSE)),
  xlab = "",
  trace.factor = (df_data|>filter(!is.na(rt)))$first_framing,
  trace.label = "First framing",
  response = (df_data|>filter(!is.na(rt)))$rt,
  ylab = "Response time (ms)",
  col = c("pink", "blue"),
  lty = 1,
  lwd = 2,
  xaxt = "n"
)
mtext("Time", side = 1, line = 1)

interaction.plot(
  x.factor = (df_data|>filter(!is.na(rt)))$trial_nr,
  xlab = "",
  trace.factor = (df_data|>filter(!is.na(rt)))$first_framing,
  trace.label = "First framing",
  response = (df_data|>filter(!is.na(rt)))$rt,
  ylab = "Response time (ms)",
  col = c("pink", "blue"),
  lty = 1,
  lwd = 2,
  xaxt = "n"
)
mtext("Time", side = 1, line = 1)

```

## Bayesian models

```{r}
# Uncomment to run, recommended to run chunk below instead
# options(mc.cores=4) # -> be carefull with adjusting this setting
# fit_acc_bayesian <- brm(
#   accuracy ~ framing * outcome * correct_response + first_framing * framing +
#   (framing * outcome * correct_response|nr),
#   data = df_data,
#   family = bernoulli(),
#   chains = 4,
#   iter = 5500,
#   warmup = 500,
#   seed = 1
# )

```

```{r}
download.file(
  "https://media.githubusercontent.com/media/JelleGoetschalckx/RPEP/main/models/fit_acc_bayesian",
  destfile = "fit_acc_bayesian.rds",
  mode = "wb"
)
fit_acc_bayesian <- readRDS("fit_acc_bayesian.rds")

summary(fit_acc_bayesian)

```

```{r}
# Uncomment to run, recommended to run chunk below instead
# options(mc.cores=4) # -> be carefull with adjusting this setting
# fit_rt_bayesian <- brm(
#   rt ~ framing * outcome + (framing * first_framing) +
#   (framing * outcome | nr),
#   data = df_data |> filter(correct_response == "Go") |> filter(!is.na(rt)),
#   chains = 4,
#   family = shifted_lognormal(),
#   iter = 5500,
#   warmup = 500,
#   seed = 1
# )

```

```{r}
download.file(
  "https://media.githubusercontent.com/media/JelleGoetschalckx/RPEP/main/models/fit_rt_bayesian",
  destfile = "fit_rt_bayesian.rds",
  mode = "wb"
)
fit_rt_bayesian <- readRDS("fit_rt_bayesian.rds")

summary(fit_rt_bayesian)

```

### Hypothesis 1: Replication of Pavlovian Biases in the Congruent Framing

1.1 G-R trials will yield more accurate responses than G-L trials.

```{r}
# Accuracy G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# correct_response            1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response    1           1           | 0
# framing:outcome             1          -1           | 2
# correct_response:outcome    1          -1           | 2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   1          -1           | 2


coef(fit_rt_bayesian)
L_h1.1_bayesian <- "
  2 * outcome1  + 
  2 * framing1:outcome1 + 
  2 * outcome1:correct_response1 + 
  2 * framing1:outcome1:correct_response1 
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h1.1_bayesian)

# Plot
recolor_plot(plot_acc_congruent, c("red", "red", "white", "white"))

# Result: proof for Pavlovian bias

```

1.2 G-R trials will yield faster response times than G-Ltrials.

```{r}
# RT G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:outcome             1          -1           | 2
# framing:first_framing       0           0           | 0

L_h1.2_bayesian <- "
  2 * outcome1 +
  2 * framing1:outcome1
  = 0
"

# Hypothesis test
brms::hypothesis(fit_rt_bayesian, L_h1.2_bayesian)

# Plot
combine_plots(
  recolor_plot(plot_rt_congruent, c("red", "red")),
  recolor_plot(plot_rt_incongruent, c("white", "white")),
  type = "rt"
)

# Result: proof for Pavlovian bias

```

1.3 NG-L trials will yield more accurate responses than NG-R trials.

```{r}
# Accuracy NG-R vs. NoGo L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                     1           1           | 0
# correct_response           -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response   -1          -1           | 0
# framing:outcome             1          -1           | 2
# correct_response:outcome   -1           1           |-2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -1           1           |-2

L_h1.3_bayesian <- "
  2 * outcome1 + 
  2 * framing1:outcome1 + 
  -2 * outcome1:correct_response1 + 
  -2 * framing1:outcome1:correct_response1 
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h1.3_bayesian)

# Plot
recolor_plot(plot_acc_congruent, c("white", "white", "red", "red"))

# Result: no proof for Pavlovian bias

```

### Hypothesis 2: Collapse of Pavlovian Bias in the Incongruent Framing

2.1 G-R trials will yield equally or less accurate responses compared to G-L trials.

```{r}
# Accuracy G-R vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# correct_response            1           1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response   -1          -1           | 0
# framing:outcome            -1           1           |-2
# correct_response:outcome    1          -1           | 2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -1           1           |-2

L_h2.1_bayesian <- "
  2 * outcome1  + 
  -2 * framing1:outcome1 + 
  2 * outcome1:correct_response1 + 
  -2 * framing1:outcome1:correct_response1 
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h2.1_bayesian)

# Plot
recolor_plot(plot_acc_incongruent, c("red", "red", "white", "white"))

# Result: proof for collapse of Pavlovian bias

```

2.2 G-R trials will yield equally or slower response times compared to G-L trials.

```{r}
# RT Go-T vs. G-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:outcome            -1           1           |-2
# framing:first_framing       0           0           | 0

L_h2.2_bayesian <- "
  2 * outcome1 +
  -2 * framing1:outcome1
  = 0
"

# Hypothesis test
brms::hypothesis(fit_rt_bayesian, L_h2.2_bayesian)

# Plot
combine_plots(
  recolor_plot(plot_rt_congruent, c("white", "white")),
  recolor_plot(plot_rt_incongruent, c("red", "red")),
  type = "rt"
)

# Result: proof for collapse of Pavlovian bias

```

2.3 NG-R trials will yield equally or less accurate responses compared to NG-L trials.

```{r}
# Accuracy NG-R vs. NG-L
#                             Reward      Loss        | Difference
# intercept                   0           0           | 0
# framing                    -1          -1           | 0
# correct_response           -1          -1           | 0
# outcome                     1          -1           | 2
# first_framing               0           0           | 0
# framing:correct_response    1           1           | 0
# framing:outcome            -1           1           |-2
# correct_response:outcome   -1           1           |-2
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   1          -1           | 2

L_h2.3_bayesian <- "
  2 * outcome1  + 
  -2 * framing1:outcome1 + 
  -2 * outcome1:correct_response1 + 
  2 * framing1:outcome1:correct_response1 
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h2.3_bayesian)

# Plot
recolor_plot(plot_acc_incongruent, c("white", "white", "red", "red"))

# Result: proof for opposite effect to Pavlovian bias

```

### Hypothesis 3: Significant difference in Pavlovian bias effects

3.1 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for accuracy in G-R vs. G-L trials.

```{r}
# Accuracy Congruent vs. Incongruent for G-R vs. G-L 
# (L_h1.1 - L_h2.1)
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# correct_response            0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:correct_response    0           0           | 0
# framing:outcome             2          -2           | 4
# correct_response:outcome    2           2           | 0
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome   2          -2           | 4

L_h3.1_bayesian <- "
  4 * framing1:outcome1  + 
  4 * framing1:outcome1:correct_response1
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h3.1_bayesian)

# Plots
combine_plots(
  recolor_plot(plot_acc_congruent, c("brown", "brown", "white", "white")),
  recolor_plot(plot_acc_incongruent, c("red", "red", "white", "white")),
  type = "acc"
)

# Result: proof for Pavlovian bias

```

3.2 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for response times in G-R vs. G-L trials.

```{r}
# RT Congruent vs. Incongruent for G-R vs. G-L
# (L_h1.2 - L_h2.2)
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:outcome             2          -2           | 4
# framing:first_framing       0           0           | 0

L_h3.2_bayesian <- "
  4 * framing1:outcome1
  = 0
"

# Hypothesis test
brms::hypothesis(fit_rt_bayesian, L_h3.2_bayesian)

# Plots
combine_plots(
  recolor_plot(plot_rt_congruent, c("brown", "brown")),
  recolor_plot(plot_rt_incongruent, c("red", "red")),
  type = "rt"
)

# Result: proof for Pavlovian bias

```

3.3 There is a difference in Pavlovian bias between the Congruent and Incongruent framing for accuracy in NG-R vs. NG-L trials.

```{r}
# Accuracy Congruent vs. Incongruent for NG-R vs. NG-L
#                             Congruent   Incongruent | Difference
# intercept                   0           0           | 0
# framing                     0           0           | 0
# correct_response            0           0           | 0
# outcome                     2           2           | 0
# first_framing               0           0           | 0
# framing:correct_response    0           0           | 0
# framing:outcome             2          -2           | 4
# correct_response:outcome   -2          -2           | 0
# framing:first_framing       0           0           | 0
# framing:corr_resp:outcome  -2           2           |-4

L_h3.3_bayesian <- "
  4 * framing1:outcome1  + 
  -4 * framing1:outcome1:correct_response1
  = 0
"

# Hypothesis test
brms::hypothesis(fit_acc_bayesian, L_h3.2_bayesian)

# Plots
combine_plots(
  recolor_plot(plot_acc_congruent, c("white", "white", "brown", "brown")),
  recolor_plot(plot_acc_incongruent, c("white", "white", "red", "red")),
  type = "acc"
)

# !!

```

### Hypothesis 4: Lasting effect of first framing

```{r}
L_h4_acc_bayesian <- "framing1:first_framing1 = 0"
brms::hypothesis(fit_acc_bayesian, L_h4_acc_bayesian)

L_h4_rt_bayesian <- "framing1:first_framing1 = 0"
brms::hypothesis(fit_rt_bayesian, L_h4_rt_bayesian)

# !!

```
